---
title: "Data 621 - Homework 5"
author: "Group 2: William Aiken, Donald Butler, Michael Ippolito, Bharani Nittala,
  and Leticia Salazar"
date: "December 11, 2022"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: yes
    toc_float: yes
  pdf_document:
    dev: cairo_pdf
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

$~$

## Overview:

We will explore, analyze and model a data set containing information on approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine. These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant. A larger wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.

## Objective

Build a count regression model to predict the number of cases of wine that will be sold given certain properties of the wine. HINT: Sometimes, the face that a variable is missing is actually predictive of the target. 

## Description

Below is a short description of the variables of interest in the data set:


| VARIABLE NAME:      | DEFINITION:                                     | THEORETICAL EFFECT:               |
|:---                 |:---:                                            |:---:                              |
|INDEX                |Identification Variable (do not use)             |None                               |
|TARGET               |Number of Cases Purchased                        |None                               |
|AcidIndex            |Proprietary method of testing totalacidity of wine by using a weighted average |     |
|Alcohol              |Alcohol Content                                  |                                   |
|Chorides             |Cholride content of wine                         |                                   |
|CitricAcid           |Citric Acid Content                              |                                   |
|Density              |Density of Wine                                  |                                   |
|FixedAcidity         |Fixed Acidity of Wiine                           |                                   |
|FreeSulfurDioxide    |Sulfur Dioxide content of wine                   |                                   |
|LabelAppeal          |Marketing Score indicating the appeal of label design for consumers. High numbers suggest customers like the label design. Negative numbers suggest customers don't like the design.   | Many consumers purchase based on the visual appeal of the wine label design. Higher numbers suggest better sales.  |
|ResidualSugar        |Residual Sugar of wine                           |                                   |
|STARS                |Wine rating by a team of experts: 4 Stars = Excellent, 1 Star = Poor |A high number of stars suggests high sales   |
|Sulphates            |Sulfate content of Wine                          |                                   |
|TotalSulfurDioxide   |Total Sulfur Dioxide of Wine                     |                                   |
|VolatileAcidity      |Volatile Acid content of wine                    |                                   |
|pH                   |pH of wine                                       |                                   |

$~$

-----------

### Load Libraries:

These are the libraries used to explore, prepare, analyze and build our models
```{r libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(corrplot)
library(skimr)
library(DataExplorer)
library(ggplot2)
library(hrbrthemes)
library(mice)
library(MASS)
library(dvmisc)
library(gridExtra)
library(lattice)
library(faraway)
library(pscl)
```

$~$

## Load Data set:

We have included the original data sets in our GitHub account and read from this location. Below we are showing the training data set:
```{r, echo=FALSE}
dftrain <- read.csv("https://raw.githubusercontent.com/letisalba/Data_621/master/Homework_5/csv/wine-training-data.csv")
dfeval <- read.csv("https://raw.githubusercontent.com/letisalba/Data_621/master/Homework_5/csv/wine-evaluation-data.csv")
head(dftrain)
```

-----------

## Data Exploration:

Using the `summary()` function lets start exploring the training and evaluation data.

Training:
```{r, echo=FALSE}
summary(dftrain)
```

$~$

Evaluation:
```{r, echo=FALSE}
summary(dfeval)
```

$~$

Using the `DataExplorer` package we use the `create_report` function which pulls a full data profile from our training data set and create an html file with basic statistics, structure, missing data, distribution visualizations, correlation matrix and principal component analysis for our data. You can find these output in our github.

```{r, eval=FALSE, message=FALSE, warning=FALSE}
# Do not render since it will produce a separate html file
# Remove TARGET from eval report since it will contain all NAs and will make the correlation plot fail to render
DataExplorer::create_report(dftrain, output_file='training_report.html')
DataExplorer::create_report(dfeval %>% select(-TARGET), output_file='eval_report.html')
```

$~$

Based on this our training data includes 12795 records and 16 variables whereas the evaluation data includes 3335 records and 16 variables.

Training:
```{r, echo=FALSE}
str(dftrain)
```

$~$

Evaluation:
```{r, echo=FALSE}
str(dfeval)
```

$~$

Lets take a look at the distribution of each variables in the training data set.

Based on the plots below, we can tell that most of the variables seem to be normally distributed with the exception of `AcidIndex` and `STARS` being right skewed. `INDEX` shows a uniform distribution but has no effect on our data so during the data preparation stage we will be removing it. 

```{r distribution, echo=FALSE, warning=FALSE, fig.height = 10, fig.width = 10, fig.align="center"}
plot_train <- dftrain %>%
  gather(key = 'variable', value = 'value')

ggplot(plot_train) +
  geom_histogram(aes(x=value, y = ..density..), bins=30) +
  geom_density(aes(x=value), color='blue') +
  theme_ipsum() +
  facet_wrap(. ~variable, scales='free', ncol=4)
```

<<<<<<< HEAD
$~$

The fact that some wines are not rated could be a potential predictor. We'll treat NAs as its own star rating.
=======
It is noted that there are a significant number of zeros in the TARGET variable, meaning that no cases were sold of that particular variety of wine. In addition, the fact that some wines are not rated could be a potential predictor. We'll treat NAs as its own star rating.

>>>>>>> 67bfcfc1ff22e55ffccd59affcbad2f3c6dec13a
```{r, echo=FALSE}
# Create logical variable to indicate whether there is a star rating for this wine
dftrain <- dftrain %>%
    mutate(STARS=ifelse(is.na(STARS), 'NR', STARS))
dfeval <- dfeval %>%
    mutate(STARS=ifelse(is.na(STARS), 'NR', STARS))
```

$~$

We'll also look at the number of cases of wine sold against the predictors.

```{r warning=FALSE, fig.width=12, fig.heigh=16, echo=FALSE}
plt <- vector('list', ncol(dftrain) - 1)
for (i in seq(3, 16)) {  # skip INDEX and TARGET variables
    if (class(dftrain[, i]) == 'numeric') {
        tmpmin <- min(dftrain[, i], na.rm=T)
        tmpinterval <- (max(dftrain[, i], na.rm=T) - tmpmin) / 5
        tmpcuts <- c()
        for (j in seq(1, 5)) {
            tmpcuts <- c(tmpcuts, tmpmin + (j * tmpinterval))
        }
        #dftrain$x <- dftrain[, i] %>% cut(breaks=5, ordered_result=T, right=F)
        dftrain$x <- dftrain[, i] %>% cut(breaks=tmpcuts, ordered_result=T, right=F)
    } else {
        dftrain$x <- dftrain[, i]
    }
    dftmp <- dftrain %>% group_by(x) %>% summarize(ct=sum(TARGET))
    plt[[i]] <- barchart(dftmp$ct ~ dftmp$x, horiz=F, col='darkgreen', xlab=colnames(dftrain)[i], ylab='Cases')
}
dftrain <- subset(dftrain, select=-x)  # remove temporary variable
grid.arrange(grobs=plt[3:7], ncol=3, nrow=2)
grid.arrange(grobs=plt[8:13], ncol=3, nrow=2)
grid.arrange(grobs=plt[14:16], ncol=3, nrow=2)
```


As shown, more cases of wine are sold for mid-range values of all categories of acidity, sugar, chlorides, the dioxides, density, pH, sulphates, and alcohol. Surprisingly, more cases were sold for labels that had mid-range label appeal. A lower acid index seemed to indicate more cases sold. And more cases were sold for wines rated only two stars, indicating that consumers may consider higher-starred wines as too pricey.


$~$

## Data Preparation:

Data preparation was performed on both the training and evaluation data sets but will only be displayed for the training data. We'll also need to removing the `INDEX` variable.
```{r, echo=FALSE}
# For some reason R renamed the INDEX column to "ï..INDEX"
dftrain <- dftrain %>% 
  dplyr::select(-INDEX)
```

```{r, echo=FALSE}
dfeval <- dfeval %>% 
  dplyr::select(-IN)
```

Now we'll impute missing values using R's Multiple Imputation by Chained Equations (MICE) package. We'll avoid imputing the STARS variable as the absence of a star rating may be a significant predictor.
```{r, echo=FALSE}
# Impute missing values in training data (except for STARS)
dftrain_imputed <- mice(dftrain %>% dplyr::select(-STARS), m=5, maxit=5, method='pmm')
cleandf <- complete(dftrain_imputed) %>%
    mutate(STARS = dftrain$STARS)

# Impute missing values in eval data (except for STARS and TARGET)
dfeval_imputed <- mice(dfeval %>% dplyr::select(-STARS, -TARGET), m=5, maxit=5, method='pmm')
cleandf_eval <- complete(dfeval_imputed) %>%
    mutate(STARS = dfeval$STARS, TARGET = dfeval$TARGET)
```

$~$

Lets look at another summary to make sure there aren't any NAs where we're not expecting them.

Training data:
```{r, echo=FALSE}
summary(cleandf)
```

$~$

Evaluation data:
```{r, echo=FALSE}
summary(cleandf_eval)
```


$~$

## Build Models:

Based on the data, we'll try two model types: a poisson general linear model and a Gaussian multiple linear model.

$~$

### Poisson Models:

* Possion Model 1
```{r, echo=FALSE}
p_mod1 <- glm(TARGET ~., family="poisson", data=cleandf)
summary(p_mod1)
```

* Possion Model with stepwise AIC approach
```{r, echo=FALSE}
p_mod2 <- stepAIC(p_mod1, trace = F)
summary(p_mod2)
```


### Multiple Linear Regression Models:

* MLR Model 1
```{r, echo=FALSE}
lm_mod1 <- lm(TARGET ~., data = cleandf)
aic_lm_mod1 = AIC(lm_mod1)
summary(lm_mod1)
```

* MLR Model 2
```{r, echo=FALSE}
lm_mod2 <- stepAIC(lm_mod1, trace = F)
aic_lm_mod2 = AIC(lm_mod2)
summary(lm_mod2)
```

$~$

### Negative Binomial Model

Since this is count data, we'll also try out a negative binomial model.

* NB model 1
```{r echo=FALSE}
nb_mod1 <- glm.nb(TARGET ~., data=cleandf)
aic_nb_mod1 = AIC(nb_mod1)
summary(nb_mod1)
```
* NB model 2
```{r echo=FALSE, warning=FALSE}
nb_mod2 <- stepAIC(nb_mod1, trace = F)
aic_nb_mod2 = AIC(nb_mod2)
summary(nb_mod2)
```

### Zero-inflated Poisson Model

Since there are a number of zeros in the target variable, we'll try a zero-inflated poisson model.

```{r}
zi_mod1 <- zeroinfl(TARGET ~ ., data=cleandf)
summary(zi_mod1)
aic_zi_mod1 = AIC(zi_mod1)
zi_mod2 <- stepAIC(zi_mod1, trace=F)
summary(zi_mod2)
aic_zi_mod2 = AIC(zi_mod2)
```

$~$


## Select Models:

In this section, an optimal model will be selected based on its performance when trained on the data. To select the models, we'll use AIC and MSE to measure accuracy of the predicted values.

Below, the Poisson and Multiple Linear Regression models have been compared to select the model with the lowest AIC.

### Comparison of Poisson Models:

We'll need to compare the AIC's of each Poisson Model.

Poisson Model 1:
```{r, echo=FALSE}
aic_p_mod1 <- p_mod1$aic
aic_p_mod1
```

Poisson Model 2:
```{r, echo=FALSE}
aic_p_mod2 <- p_mod2$aic
aic_p_mod2
```

$~$

Poisson Model 2 proves to have the lower AIC of the two, with a 33947.74 AIC. Below is the formula for Poisson Model 2.
```{r, echo=FALSE}
# Poisson - Minimum AIC
c(p_mod1$formula,p_mod2$formula)[which.min(c(p_mod1$aic,p_mod2$aic))]
```

$~$

### Comparsion of Multiple Linear Models:

We'll need to compare the Adjusted R Squares of each Linear Model.

Linear Model 1:
```{r, echo=FALSE}
r2_lm_mod1 <- summary(lm_mod1)$adj.r.squared
r2_lm_mod1
```

Linear Model 2:
```{r, echo=FALSE}
r2_lm_mod2 <- summary(lm_mod2)$adj.r.squared
r2_lm_mod2
```

$~$

Linear Model 2 proves to have the higher Adjusted R Squares, with a value of 0.540473. Below is the formula for Linear Model 2.
```{r, echo=FALSE}
# Multiple Linear Regression Model - Highest Adjusted R Squared
c(formula(lm_mod1),formula(lm_mod2))[which.max(c(summary(lm_mod1)$adj.r.squared, summary(lm_mod2)$adj.r.squared))]
```

$~$

### Comparsion of Negative Binomial Models:

Now we'll compare the AICs of the two negative binomial models.

NB Model 1:
```{r, echo=FALSE}
aic_nb_mod1
```

NB Model 2:
```{r, echo=FALSE}
aic_nb_mod2
```

The AIC of model 2 is somewhat lower than that of model 1.

$~$

### Comparsion of Zero-inflated Poisson Models:

Now we'll compare the AICs of the two zero-inflated binomial models.

NB Model 1:
```{r, echo=FALSE}
aic_zi_mod1
```

NB Model 2:
```{r, echo=FALSE}
aic_zi_mod2
```

The AIC of model 2 is somewhat lower than that of model 1.

$~$

### Mean Square Error:

The Mean Square Error measures the averaged square different between the estimated values and the actual value. The lower the value of the MSE, the more accurately the model is able to predict the values.

$$\large \text{MSE} = \large \frac{1}{n} \sum(y - \hat{y})^2$$

```{r, echo=FALSE}
mse <- function(df, model){
  mean((df$TARGET - predict(model))^2)
}
```


```{r, echo=FALSE, warning=FALSE}
mse_p_mod1 <- mse(cleandf, p_mod1)
mse_p_mod2 <- mse(cleandf, p_mod2)
mse_lm_mod1 <- get_mse(lm_mod1)
mse_lm_mod2 <- get_mse(lm_mod2)
mse_nb_mod1 <- mse(cleandf, nb_mod1)
mse_nb_mod2 <- mse(cleandf, nb_mod2)
mse_zi_mod1 <- mse(cleandf, zi_mod1)
mse_zi_mod2 <- mse(cleandf, zi_mod2)
```

$~$

### Comparison of Models:

By evaluating the AIC's and MSE's of each model, we can choose the best one be looking at the lowest AIC and lowest MSE.

```{r, echo=FALSE}
models <- c("Possion Model 1", "Possion Model 2", "Linear Model 1", "Linear Model 2", 
            "Neg Binom Model 1", "Neg Binom Model 2", "Zero-Infl Model 1", "Zero-Infl Model 2")
#rows <- c("Models", "MSE", "AIC")
MSE <- round(c(mse_p_mod1, mse_p_mod2, mse_lm_mod1, mse_lm_mod2, mse_nb_mod1, mse_nb_mod2, mse_zi_mod1, mse_zi_mod2), 1)
AIC <- round(c(aic_p_mod1, aic_p_mod2, aic_lm_mod1, aic_lm_mod2, aic_nb_mod1, aic_nb_mod2, aic_zi_mod1, aic_zi_mod2), 1)
knitr::kable(rbind(MSE, AIC), col.names = models)
```

$~$

Based on the above, the linear model has better model statistics than the poisson model.

$~$

While the linear model may have better statistics, we'll need to validate whether the assumptions of the linear model hold:
1) Homoschedastic residuals
2) Normally distributed residuals
3) Independence of predictors
4) Linearity of response

```{r, fig.width=12, echo=FALSE, warning=FALSE}

# Load libraries
library(car)
library(lmtest)
library(olsrr)

# Define function to calculate mean squared error
calc_mse <- function(lmod) {
  return(mean((summary(lmod))$residuals ^ 2))
}

# Define function to aid in model analysis
ModelAnalysis <- function(lmod) {

  # Plot residuals
  print('--------------------------------------------------')
  print(lmod$call)
  par(mfrow=c(2,2))
  plot(lmod)
  print('')

  # Shapiro test to determine normality of residuals
  # Null hypothesis: the residuals are normal.
  # If the p-value is small, reject the null, i.e., consider the residuals *not* normally distributed.
  if (length(lmod$fitted.values) > 3 & length(lmod$fitted.values) < 5000) {
      st <- shapiro.test(lmod$residuals)
      if (st$p.value <= 0.05) {
        print(paste0("Shapiro test for normality: The p-value of ", st$p.value, " is <= 0.05, so reject the null; i.e., the residuals are NOT NORMAL"))
      } else {
        print(paste0("Shapiro test for normality: The p-value of ", st$p.value, " is > 0.05, so do not reject the null; i.e., the residuals are NORMAL"))
      }
      print('')
  } else {
      print("Shapiro test for normality of residuals cannot be performed; sample length must be between 3 and 5000.")
  }
     
  # Breusch-Pagan test to determine homoschedasticity of residuals
  # Null hypothesis: the residuals are homoschedastic.
  # If the p-value is small, reject the null, i.e., consider the residuals heteroschedastic.
  bp <- bptest(lmod)
  if (bp$p.value > 0.05 & bp$statistic < 10) {
      print(paste0("Breusch-Pagan test for homoschedasticity: The p-value of ", bp$p.value, " is > 0.05 and the test statistic of ", bp$statistic,
          " is < 10, so don't reject the null; i.e., the residuals are HOMOSCHEDASTIC"))
  } else if (bp$p.value <= 0.05) {
      print(paste0("Breusch-Pagan test for homoschedasticity: The p-value of ", bp$p.value, " is <= 0.05 and the test statistic is ", bp$statistic,
          ", so reject the null; i.e., the residuals are HETEROSCHEDASTIC"))
  } else {
      print(paste0("Breusch-Pagan test for homoschedasticity: The p-value of ", bp$p.value, " and test statistic of ", bp$statistic,
          " are inconclusive, so homoschedasticity can't be determined using this test."))
  }
  print('')

  # Visually look for colinearity - dont do this for large models
  #pairs(model.matrix(lmod))

  # Variance inflation factor (VIF)
  print('Variance inflation factor (VIF)')
  print('<=1: not correlated, 1-5: moderately correlated, >5: strongly correlated')
  print(sort(vif(lmod), decreasing=T))
  print('')
  
  # Standardized residual plots (look for points outside of 2 or 3 stdev)
  p <- length(summary(lmod)$coeff[,1] - 1)  # number of model parameters
  stanres <- rstandard(lmod)
  for (i in seq(1, ceiling(p / 4))) {
    par(mfrow=c(2,2))
    starti <- ((i - 1) * 4) + 1
    for (j in seq(starti, starti + 3)) {
      if (j + 1 <= ncol(model.matrix(lmod))) {
        # Skip these plots since we're pretty sure that a linear model isn't valid here
        #plot(model.matrix(lmod)[, j + 1], stanres, xlab=colnames(model.matrix(lmod))[j + 1], ylab='Standardized residuals')
        #abline(h=c(-2, 2), lt=3, col='blue')
        #abline(h=c(-3, 3), lt=2, col='red')
      }
    }
  }
  
  # Model scores
  print('Model scores:')
  print(paste0('    adjusted R-squared: ', round(summary(lmod)$adj.r.squared, 3)))
  print(paste0('    AIC: ', round(AIC(lmod, k=2), 3)))
  print(paste0('    BIC: ', round(BIC(lmod), 3)))
  print(paste0('    Mallow\'s Cp: ', round(ols_mallows_cp(lmod, fullmodel=lmod), 3)))
  m <- mse(lmod)
  print(paste0('    mean squared error: ', round(calc_mse(lmod), 3)))
  print('')
  
  # Find leverage point cutoff
  n <- length(lmod$residuals)
  cutoff <- 2 * (p + 1) / n
  print(paste0('Leverage point cutoff: ', cutoff))
  print('')

  # Show points of influence
  print('First 10 points of influence:')
  poi <- lm.influence(lmod)$hat
  len_poi <- length(poi)
  ct <- 0
  for (i in seq(1, length(poi))) {
    if (poi[i] > cutoff) {
      ct <- ct + 1
      print(paste0('    case #', i, ': ', round(poi[i], 3)))
    }
        if (ct > 10) {
            break
        }
  }
  print('')
  
}

```

```{r echo=FALSE}
ModelAnalysis(lm_mod1)
ModelAnalysis(lm_mod2)
```

As shown, the residuals are heteroschedastic and exhibit a clearly defined pattern. While the Shapiro test for normality couldn't be performed due to the sample size, the QQ plot shows visually that the residuals are not normally distributed.

Now we'll see if the Poisson models exhibit any overdispersion, which could lead us to invalidate the models and go with the negative binomial model instead. This can be done by dividing the residual deviance by the degrees of freedom.

```{r echo=FALSE}
op1 <- p_mod1$deviance / p_mod1$df.residual
op2 <- p_mod2$deviance / p_mod2$df.residual
print(paste0('Poisson model 1 overdispersion: ', op1))
print(paste0('Poisson model 2 overdispersion: ', op2))
```

Since the overdispersion parameter isn't much greater than 1 (generall, less than 1.10), this suggests that the poisson model is a good fit, and we can be relatively confident that we don't need to go with the negative binomial.

Now we'll check the model accuracy by predicting the number of cases sold based on our model parameters. Since we're dealing with whole cases of wine, we'll round the prediction to the nearest integer.

```{r echo=FALSE, warning=FALSE}

# Negative binomial
cleandfnew <- cleandf
cleandfnew$pred_target <- exp(predict(nb_mod2, cleandfnew, interval ='prediction'))
cleandfnew$correct <- ifelse(round(cleandfnew$pred_target, 0) == cleandfnew$TARGET, 1, 0)
acc_num <- sum(cleandfnew$correct)
acc_pct <- round(100 * acc_num / nrow(cleandfnew), 1)
print(paste0('Negative binomial model accuracy: ', acc_num, ' of ', nrow(cleandfnew), ' (', acc_pct, '%)'))
hist(cleandfnew$TARGET, xlab='Cases', main='Actual Cases Sold')
hist(round(cleandfnew$pred_target, 0), xlab='Cases', main='Predicted Cases Sold')

# Poisson
cleandfnew$pred_target <- exp(predict(p_mod2, cleandfnew, interval ='prediction'))
cleandfnew$correct <- ifelse(round(cleandfnew$pred_target, 0) == cleandfnew$TARGET, 1, 0)
acc_num <- sum(cleandfnew$correct)
acc_pct <- round(100 * acc_num / nrow(cleandfnew), 1)
print(paste0('Poisson accuracy: ', acc_num, ' of ', nrow(cleandfnew), ' (', acc_pct, '%)'))
hist(cleandfnew$TARGET, xlab='Cases', main='Actual Cases Sold')
hist(round(cleandfnew$pred_target, 0), xlab='Cases', main='Predicted Cases Sold')

# Zero-inflated
cleandfnew$pred_target <- predict(zi_mod2, cleandfnew, interval ='prediction')
cleandfnew$correct <- ifelse(round(cleandfnew$pred_target, 0) == cleandfnew$TARGET, 1, 0)
acc_num <- sum(cleandfnew$correct)
acc_pct <- round(100 * acc_num / nrow(cleandfnew), 1)
print(paste0('Zero-inflated poisson accuracy: ', acc_num, ' of ', nrow(cleandfnew), ' (', acc_pct, '%)'))
hist(round(cleandfnew$pred_target, 0), xlab='Cases', main='Predicted Cases Sold')

```

None of the models performs very well. Because of the zero counts, we'll take a different tack rather than a simple single-model approach. Instead, we'll try doing a hierarchical model, in which we'll first model zero counts using a binomial model, then we'll model positive counts with a different count-based model.

```{r echo=FALSE}
# Binomial on whether cases were sold or not
cleandf_binom <- cleandf %>%
    mutate(pos_cases=ifelse(TARGET > 0, T, F))
b_mod1 <- glm(pos_cases ~ . - TARGET, family=binomial(), data=cleandf_binom)
b_mod2 <- stepAIC(b_mod1, trace=F)
aic_b_mod1 <- AIC(b_mod1)
aic_b_mod2 <- AIC(b_mod2)
```

Summary of binomial model 1:
```{r echo=FALSE}
summary(b_mod1)
```

Summary of binomial model 2:
```{r echo=FALSE}
summary(b_mod2)
```

AIC of binomial model 1:
```{r echo=FALSE}
aic_b_mod1
```

AIC of binomial model 2:
```{r echo=FALSE}
aic_b_mod2
```

Since the AIC of the second model is lower, we'll use it to predict whether any cases of each wine variety were sold. Then we'll check the accuracy of the predictions against the training data.

```{r echo=FALSE}
cleandf_binom$pred_p <- ilogit(predict(b_mod2, cleandf_binom, interval='prediction'))
cleandf_binom$pred_pos_cases <- ifelse(cleandf_binom$pred_p > 0.5, T, F)
cleandf_binom$correct_pos_cases <- ifelse(cleandf_binom$pred_pos_cases == cleandf_binom$pos_cases, 1, 0)
acc_num <- sum(cleandf_binom$correct_pos_cases)
acc_pct <- round(100 * acc_num / nrow(cleandf_binom), 1)
print(paste0('Binomial model accuracy with p of 0.5: ', acc_num, ' of ', nrow(cleandf_binom), ' (', acc_pct, '%)'))
```

```{r echo=FALSE}
roc_func <- function(data){
  temp_x <- rep(0, 101)
  temp_y <- rep(0, 101)
  temp_z <- rep(0, 101)
  temp_seq <- seq(from = 0, to = 1, by = 0.01)
  max_acc <- 0
  p_max_acc <- 0
  for (i in 1:length(temp_seq)){
    df <- data %>% 
        mutate(scored.class = as.logical(pred_p > temp_seq[i])) %>%
        mutate(true.class = as.logical(TARGET > 0)) %>%
        mutate(
           TP = ifelse(true.class == T & scored.class == T, 1, 0),
           FP = ifelse(true.class == F & scored.class == T, 1, 0),
           FN = ifelse(true.class == T & scored.class == F, 1, 0),
           TN = ifelse(true.class == F & scored.class == F, 1, 0)
        )
    TPR <- sum(df$TP)/(sum(df$TP) + sum(df$FN))
    FPR <- sum(df$FP)/(sum(df$FP) + sum(df$TN))
    acc <- (sum(df$TP) + sum(df$TN)) / nrow(df)
    if (acc > max_acc) {
        max_acc <- acc
        p_max_acc <- temp_seq[i]
    }
    temp_x[i] <- FPR
    temp_y[i] <- TPR
    temp_z[i] <- acc
  }
  temp_df <- bind_cols(temp_x, temp_y, temp_seq, temp_z) %>% as.data.frame()
  names(temp_df) <- c("FPR", "TPR", "p", "Accuracy")
  plt <- ggplot2::ggplot(data = temp_df, aes(x = FPR, y = TPR)) + geom_point() + geom_abline()
  plt2 <- ggplot2::ggplot(data = temp_df, aes(x = p, y = Accuracy)) + geom_point()
  AUC <- pracma::trapz(temp_x, temp_y)
  output <- list(plt, plt2, AUC, max_acc, p_max_acc)
  return(output)
}

# Generate ROC curve
rf <- roc_func(cleandf_binom)
print(rf)
rf[[5]]

# 
cleandf_binom$pred_p <- ilogit(predict(b_mod2, cleandf_binom, interval='prediction'))
cleandf_binom$pred_pos_cases <- ifelse(cleandf_binom$pred_p > rf[[5]], T, F)
cleandf_binom$correct_pos_cases <- ifelse(cleandf_binom$pred_pos_cases == cleandf_binom$pos_cases, 1, 0)
acc_num <- sum(cleandf_binom$correct_pos_cases)
acc_pct <- round(100 * acc_num / nrow(cleandf_binom), 1)
print(paste0('Binomial model accuracy with p of ', rf[[5]], ': ', acc_num, ' of ', nrow(cleandf_binom), ' (', acc_pct, '%)'))

```

Since the accuracy is very good, we'll use that to predict zero counts, then we'll use a separate count-based model to predict positive case counts.

Start with poisson modeling:
```{r echo=FALSE}
p_mod3 <- glm(TARGET ~ ., family=poisson(), data=cleandf['TARGET' > 0])
aic_p_mod3 <- AIC(p_mod3)
p_mod4 <- stepAIC(p_mod3, trace=F)
aic_p_mod4 <- AIC(p_mod4)
print(paste0('AIC of poisson full model: ', aic_p_mod3))
print(paste0('AIC of poisson step-reduced model: ', aic_p_mod4))
```

Check for overdispersion of the poisson models:
```{r echo=FALSE}
op3 <- p_mod3$deviance / p_mod3$df.residual
op4 <- p_mod4$deviance / p_mod4$df.residual
print(paste0('Poisson full model overdispersion: ', op3))
print(paste0('Poisson step-reduced model overdispersion: ', op4))
```

Neither of these is much greater than one, but we'll try negative binomial modeling to see if it yields a better result anyway.
```{r echo=FALSE, warning=FALSE}
nb_mod3 <- glm.nb(TARGET ~ ., data=cleandf['TARGET' > 0])
aic_nb_mod3 <- AIC(nb_mod3)
nb_mod4 <- stepAIC(nb_mod3, trace=F)
aic_nb_mod4 <- AIC(nb_mod4)
print(paste0('AIC of negative binomial full model: ', aic_nb_mod3))
print(paste0('AIC of negative binomial step-reduced model: ', aic_nb_mod4))
```

Since the AIC of the poisson step-reduced model is lowest, we'll use it to predict cases sold against the training data.

```{r echo=FALSE}

# Vary p
tmp_p <- seq(from = 0, to = 1, by = 0.01)
c_acc_pct <- c()
max_acc <- 0
p_max_acc <- 0
for (i in seq(1, length(tmp_p))) {
    # Make initial prediction on zero cases vs non-zero cases
    cleandf$pred_p <- ilogit(predict(b_mod2, cleandf, interval='prediction'))
    cleandf$pred_target <- ifelse(cleandf$pred_p > tmp_p[i], NA, 0)
    # Split df into zero cases and non-zero cases
    cleandf1 <- cleandf %>%
        filter(pred_target == 0)
    cleandf2 <- cleandf %>%
        filter(is.na(pred_target))
    cleandf2$pred_target <- round(exp(predict(p_mod4, cleandf2, interval ='prediction')), 0)
    cleandfnew <- rbind(cleandf1, cleandf2)
    cleandfnew$correct <- ifelse(cleandfnew$pred_target == cleandfnew$TARGET, 1, 0)
    acc_num <- sum(cleandfnew$correct)
    acc_pct <- round(100 * acc_num / nrow(cleandfnew), 1)
    c_acc_pct <- c(c_acc_pct, acc_pct)
    if (acc_pct > max_acc) {
        max_acc <- acc_pct
        p_max_acc <- tmp_p[i]
    }
    #print(paste0('Hierarchical model accuracy: ', acc_num, ' of ', nrow(cleandfnew), ' (', acc_pct, '%)'))
}
plot(c_acc_pct ~ tmp_p)
print(paste0('Max accuracy is ', max_acc, ' when p=', p_max_acc))

```





Prediction from optimal model:
```{r, echo=FALSE}
prob2 <- exp(predict(p_mod2, cleandf_eval, interval ='prediction'))
cleandf_eval$TARGET <- prob2
cleandf_eval %>% head(10) %>% as_tibble()
write.csv(cleandf_eval, "wine_predictions2.csv", row.names = FALSE)
```

$~$

## Appendix:

```{r, eval=FALSE}
# load libaries
library(tidyverse)
library(dplyr)
library(corrplot)
library(skimr)
library(DataExplorer)
library(ggplot2)
library(hrbrthemes)
library(mice)

# load data
dftrain <- read.csv("https://raw.githubusercontent.com/letisalba/Data_621/master/Homework_5/csv/wine-training-data.csv")
dfeval <- read.csv("https://raw.githubusercontent.com/letisalba/Data_621/master/Homework_5/csv/wine-evaluation-data.csv")
head(dftrain)

# summary of training and evaluation data sets
summary(dftrain)
summary(dfeval)

# Do not render since it will produce a separate html file
# Remove TARGET from eval report since it will contain all NAs and will make the correlation plot fail to render
DataExplorer::create_report(dftrain, output_file='training_report.html')
DataExplorer::create_report(dfeval %>% select(-TARGET), output_file='eval_report.html')

# structure of training and evaluation data 
str(dftrain)
str(dfeval)

# plotting distribution of training data 
plot_train <- dftrain %>%
  gather(key = 'variable', value = 'value')

ggplot(plot_train) +
  geom_histogram(aes(x=value, y = ..density..), bins=30) +
  geom_density(aes(x=value), color='blue') +
  theme_ipsum() +
  facet_wrap(. ~variable, scales='free', ncol=4)

# Create logical variable to indicate whether there is a star rating for this wine
dftrain <- dftrain %>%
    mutate(STARS=ifelse(is.na(STARS), 'NR', STARS))
dfeval <- dfeval %>%
    mutate(STARS=ifelse(is.na(STARS), 'NR', STARS))

#Look at the number of cases of wine sold against the predictors.
plt <- vector('list', ncol(dftrain) - 1)
for (i in seq(3, 16)) {  # skip INDEX and TARGET variables
    if (class(dftrain[, i]) == 'numeric') {
        tmpmin <- min(dftrain[, i], na.rm=T)
        tmpinterval <- (max(dftrain[, i], na.rm=T) - tmpmin) / 5
        tmpcuts <- c()
        for (j in seq(1, 5)) {
            tmpcuts <- c(tmpcuts, tmpmin + (j * tmpinterval))
        }
        #dftrain$x <- dftrain[, i] %>% cut(breaks=5, ordered_result=T, right=F)
        dftrain$x <- dftrain[, i] %>% cut(breaks=tmpcuts, ordered_result=T, right=F)
    } else {
        dftrain$x <- dftrain[, i]
    }
    dftmp <- dftrain %>% group_by(x) %>% summarize(ct=sum(TARGET))
    plt[[i]] <- barchart(dftmp$ct ~ dftmp$x, horiz=F, col='darkgreen', xlab=colnames(dftrain)[i], ylab='Cases')
}
dftrain <- subset(dftrain, select=-x)  # remove temporary variable
grid.arrange(grobs=plt[3:7], ncol=3, nrow=2)
grid.arrange(grobs=plt[8:13], ncol=3, nrow=2)
grid.arrange(grobs=plt[14:16], ncol=3, nrow=2)

# Removing INDEX from training and eval data 
# For some reason R renamed the INDEX column to "ï..INDEX"
dftrain <- dftrain %>% 
  dplyr::select(-ï..INDEX)
dfeval <- dfeval %>% 
  dplyr::select(-IN)

# Impute missing values in training data
dftrain_imputed <- mice(dftrain, m=5, maxit=5, method='pmm')
cleandf <- complete(dftrain_imputed) %>%
    mutate(STARS = dftrain$STARS)

# Impute missing values in eval data (except for TARGET)
dfeval_imputed <- mice(dfeval %>% select(-TARGET), m=5, maxit=5, method='pmm')
cleandf_eval <- complete(dfeval_imputed) %>%
    mutate(STARS = dfeval$STARS, TARGET = dfeval$TARGET)

# Look at another summary to make sure there aren't any NAs where we're not expecting them
summary(cleandf)
summary(cleandf_eval)

# Poisson model
p_mod1 <- glm(TARGET ~., family="poisson", data=cleandf)
summary(p_mod1)

# Possion Model with stepwise AIC approach
p_mod2 <- stepAIC(p_mod1, trace = F)
summary(p_mod2)

# Multiple Linear Regression Models:

# MLR Model 1
lm_mod1 <- lm(TARGET ~., data = cleandf)
aic_lm_mod1 = AIC(lm_mod1)
summary(lm_mod1)

# MLR Model 2
lm_mod2 <- stepAIC(lm_mod1, trace = F)
aic_lm_mod2 = AIC(lm_mod2)
summary(lm_mod2)

# NB model 1
nb_mod1 <- glm.nb(TARGET ~., data=cleandf)
aic_nb_mod1 = AIC(nb_mod1)
summary(nb_mod1)

# NB model 2
nb_mod2 <- stepAIC(nb_mod1, trace = F)
aic_nb_mod2 = AIC(nb_mod2)
summary(nb_mod2)

# Select Models:

# Comparison of Poisson Models:

# Poisson Model 1:
aic_p_mod1 <- p_mod1$aic
aic_p_mod1

# Poisson Model 2:
aic_p_mod2 <- p_mod2$aic
aic_p_mod2

# Poisson - Minimum AIC
c(p_mod1$formula,p_mod2$formula)[which.min(c(p_mod1$aic,p_mod2$aic))]

# Comparsion of Multiple Linar Models:

# Linear Model 1:
r2_lm_mod1 <- summary(lm_mod1)$adj.r.squared
r2_lm_mod1

# Linear Model 2:
r2_lm_mod2 <- summary(lm_mod2)$adj.r.squared
r2_lm_mod2

# Multiple Linear Regression Model - Highest Adjusted R Squared
c(formula(lm_mod1),formula(lm_mod2))[which.max(c(summary(lm_mod1)$adj.r.squared, summary(lm_mod2)$adj.r.squared))]

# Mean Square Error:
mse <- function(df, model){
  mean((df$TARGET - predict(model))^2)
}
mse_p_mod1 <- mse(cleandf, p_mod1)
mse_p_mod2 <- mse(cleandf, p_mod2)
mse_lm_mod1 <- get_mse(lm_mod1)
mse_lm_mod2 <- get_mse(lm_mod2)

# Comparison of Possion and Negative Binomial Model's:
models <- c("Possion Model 1", "Possion Model 2", "Linear Model 1", "Linear Model 2")
#rows <- c("Models", "MSE", "AIC")
MSE <- list(mse_p_mod1, mse_p_mod2, mse_lm_mod1, mse_lm_mod2)
AIC <- list(aic_p_mod1, aic_p_mod2, aic_lm_mod1, aic_lm_mod2)
knitr::kable(rbind(MSE, AIC), col.names = models)

# Prediction from optimal multiple linear regression model
prob2 <- predict(lm_mod2, cleandf_eval, interval ='prediction')
cleandf_eval$TARGET <- prob2[,1]
cleandf_eval %>% head(10) %>% as_tibble()
write.csv(cleandf_eval, "wine_predictions2.csv", row.names = FALSE)

```


-----------

## References:
https://englianhu.files.wordpress.com/2016/01/faraway-extending-the-linear-model-with-r-e28093-2006.pdf
