---
title: "DATA621_HW1"
author: "William Aiken"
date: "9/20/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```

## 1. DATA EXPLORATION

Due to the number of fields in this data, I broke the dataset into intuitive sections and explored each section individually.

```{r}
eval_df <- read.csv("https://raw.githubusercontent.com/catfoodlover/DATA621/main/HW1/moneyball-evaluation-data.csv")
train_df <- read.csv("https://raw.githubusercontent.com/catfoodlover/DATA621/main/HW1/moneyball-training-data.csv")
```


### Base Hits by Batter

*  TARGET_WINS       - Number of wins
*  TEAM_BATTING_H		- Base Hits by batters (1B,2B,3B,HR)
*  TEAM_BATTING_2B		- Doubles by batters (2B)
*  TEAM_BATTING_3B		- Triples by batters (3B)
*  TEAM_BATTING_HR		- Homeruns by batters (4B)

The means and medians are very similar for the base hits variables implying little skew to the distributions.

```{r, warning=FALSE}
train_df %>% select(c("TARGET_WINS", "TEAM_BATTING_H","TEAM_BATTING_2B","TEAM_BATTING_3B","TEAM_BATTING_HR")) %>% gtsummary::tbl_summary(statistic =list(c("TARGET_WINS", "TEAM_BATTING_H","TEAM_BATTING_2B","TEAM_BATTING_3B","TEAM_BATTING_HR") ~ "{mean} {median} {sd}"
))
```


We see tight distributions except for all base hits by batters (TEAM_BATTING_H).


```{r, message=FALSE, warning=FALSE}
temp <- train_df %>% select(c("TARGET_WINS", "TEAM_BATTING_H","TEAM_BATTING_2B","TEAM_BATTING_3B","TEAM_BATTING_HR"))

ggplot2::ggplot(stack(temp), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title = "Base Hit Variables")
```

Unsurprisingly, all possible base hits (TEAM_BATTING_H) is correlated with winning.  As you increase the number of bases achieved by an at bat, the correlation decreases.

Interestingly, doubles and triples are correlated with base hits while home runs are not.


```{r, message=FALSE, warning=FALSE}
train_df %>% select(c("TARGET_WINS", "TEAM_BATTING_H","TEAM_BATTING_2B","TEAM_BATTING_3B","TEAM_BATTING_HR")) %>% GGally::ggpairs()
```

### Batting


*  TARGET_WINS       - Number of wins
*  TEAM_BATTING_BB		- Walks by batters 
*  TEAM_BATTING_HBP	- Batters hit by pitch (get a free base) 
*  TEAM_BATTING_SO		- Strikeouts by batters 
*  TEAM_BASERUN_SB		- Stolen bases 
*  TEAM_BASERUN_CS		- Caught stealing


The measures of central tendency show us that most of these variable have slight skew to their distributions. Stolen bases has a large right skew to its distribution.


We are missing values for stikeouts, stolen bases and caught stealing.


```{r}
train_df %>% select(c("TEAM_BATTING_BB", "TEAM_BATTING_SO", "TEAM_BASERUN_SB", "TEAM_BASERUN_CS")) %>% gtsummary::tbl_summary(statistic =list(c("TEAM_BATTING_BB", "TEAM_BATTING_SO", "TEAM_BASERUN_SB", "TEAM_BASERUN_CS") ~ "{mean} {median} {sd}"
))
```

```{r, message=FALSE, warning=FALSE}
temp <- train_df %>% select(c("TEAM_BATTING_BB", "TEAM_BATTING_SO", "TEAM_BASERUN_SB", "TEAM_BASERUN_CS"))

ggplot2::ggplot(stack(temp), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title = "Batting Variables")
```


Of all the batting variables, only walks by batter has a correlation to wins.


```{r, message=FALSE, warning=FALSE}
train_df %>% select(c("TARGET_WINS", "TEAM_BATTING_BB", "TEAM_BATTING_SO", "TEAM_BASERUN_SB", "TEAM_BASERUN_CS")) %>% GGally::ggpairs()
```

### Fielding

*  TARGET_WINS       - Number of wins
*  TEAM_FIELDING_E		- Errors 
*  TEAM_FIELDING_DP	- Double Plays


The Errors variable(TEAM_FIELDING_E) has an incredibly right skewed distribution. We are missing some Double Plays values.


```{r}
train_df %>% select(c("TEAM_FIELDING_E", "TEAM_FIELDING_DP")) %>% gtsummary::tbl_summary(statistic =list(c("TEAM_FIELDING_E", "TEAM_FIELDING_DP") ~ "{mean} {median} {sd}"
))
```

```{r, message=FALSE, warning=FALSE}
temp <- train_df %>% select(c("TEAM_FIELDING_E", "TEAM_FIELDING_DP"))

ggplot2::ggplot(stack(temp), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title = "Fielding Variables")
```


Both the Fielding variables are negatively correlated with Wins.


```{r, message=FALSE, warning=FALSE}
train_df %>% select(c("TARGET_WINS", "TEAM_FIELDING_E", "TEAM_FIELDING_DP"))%>% GGally::ggpairs()
```

### Pitching

*  TARGET_WINS       - Number of wins
*  TEAM_PITCHING_BB	- Walks allowed 
*  TEAM_PITCHING_H		- Hits allowed 
*  TEAM_PITCHING_HR	- Homeruns allowed 
*  TEAM_PITCHING_SO	- Strikeouts by pitchers


Hits allowed (TEAM_PITCHING_H) has a right skew and we are missing some Strikeouts by pitcher (TEAM_PITCHING_SO) values.


```{r}
train_df %>% select(c("TEAM_PITCHING_H", "TEAM_PITCHING_HR", "TEAM_PITCHING_BB", "TEAM_PITCHING_SO")) %>% gtsummary::tbl_summary(statistic =list(c("TEAM_PITCHING_H", "TEAM_PITCHING_HR", "TEAM_PITCHING_BB", "TEAM_PITCHING_SO") ~ "{mean} {median} {sd}"
))
```

```{r, message=FALSE, warning=FALSE}
temp <- train_df %>% select(c("TEAM_PITCHING_H", "TEAM_PITCHING_HR", "TEAM_PITCHING_BB", "TEAM_PITCHING_SO"))

ggplot2::ggplot(stack(temp), aes(x = ind, y = values)) +
  geom_boxplot() + labs(title = "Pitching Variables")
```

Hits allowed is negatively correlated with Winning.

Interestingly, Home runs allowed is positively correlated with Winning.


```{r, message=FALSE, warning=FALSE}
train_df %>% select(c("TARGET_WINS","TEAM_PITCHING_H", "TEAM_PITCHING_HR", "TEAM_PITCHING_BB", "TEAM_PITCHING_SO"))%>% GGally::ggpairs()
```