---
title: 'Data 621 Homework #1'
author: "Group 2: William Aiken, Donald Butler, Michael Ippolito, Bharani Nittala, Leticia Salazar and Santiago Torres"
date: "09-25-2022"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(DataExplorer)
library(ggplot2)
library(tidyr)
library(corrplot)
library(mice)
```



## Data Overview
The data set contains approximately 2276 records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. 
Each record has the performance of the team for the given year, with all of the statistics adjusted to match the performance of a 162 game season.
Below is a short description of the variables

* INDEX 				- Identification Variable
* TARGET_WINS           - Number of wins
* TEAM_BATTING_H		- Base Hits by batters (1B,2B,3B,HR)
* TEAM_BATTING_2B		- Doubles by batters (2B)
* TEAM_BATTING_3B		- Triples by batters (3B)
* TEAM_BATTING_HR		- Homeruns by batters (4B)
* TEAM_BATTING_BB		- Walks by batters 
* TEAM_BATTING_HBP	    - Batters hit by pitch (get a free base) 
* TEAM_BATTING_SO		- Strikeouts by batters 
* TEAM_BASERUN_SB		- Stolen bases 
* TEAM_BASERUN_CS		- Caught stealing 
* TEAM_FIELDING_E		- Errors 
* TEAM_FIELDING_DP	    - Double Plays 
* TEAM_PITCHING_BB	    - Walks allowed 
* TEAM_PITCHING_H		- Hits allowed 
* TEAM_PITCHING_HR	    - Homeruns allowed 
* TEAM_PITCHING_SO	    - Strikeouts by pitchers 

## Objective
To build a multiple linear regression model on the training data to predict *TARGET_WINS*, which is the number of wins for the team.


## Data Exploration

```{r read-csv}
# read data
baseball_df <- read.csv('https://raw.githubusercontent.com/letisalba/Data_621/master/Homework_1/csv/moneyball-training-data.csv')
baseball_eval <- read.csv('https://raw.githubusercontent.com/letisalba/Data_621/master/Homework_1/csv/moneyball-evaluation-data.csv')

head(baseball_df)
dim(baseball_df)
```




```{r, baseball-summary}
summary(baseball_df)
print('Number of observations:') 
nrow(baseball_df)
print('Observations per year, 1871 - 2006:') 
round(nrow(baseball_df)/(2006-1871),2)
```

Some columns have maximum values that are clearly outliers, like TEAM_PITCHING_H AND TEAM_PITCHING_HR. The assignment mentions
that some of the season records were adjusted to match the performance during a 162-game season. 
There are 2276 seasons in the training set. Observations span 128 years, with an average of 17 teams playing per year. 

```{r}
# distribution
plot_histogram(baseball_df)
```

```{r}
# against the response variable
plot_scatterplot(baseball_df, by = "TARGET_WINS")
```


```{r}
# boxplot for train dataset
plot_boxplot(baseball_df, by="TARGET_WINS")
```


```{r, corrplot}

corrplot(cor(baseball_df[,2:17], use = 'complete.obs'))


```

Looking at the correlation plot, there appear to be several strong correlations between explanatory variables and the target.  
  
From an initial inspection, it appears the team should focus on getting players on base through hits or walks.  
Teams can still win if the pitchers allow homeruns, hits and walks to the other team.
*Variables with Highest Positive Correlation with TARGET_WINS:*  
  * TEAM_BATTING_H = 0.47
  * TEAM_BATTING_HR = 0.42
  * TEAM_BATTING_BB = 0.47
  * TEAM_PITCHING_H = 0.47 
  * TEAM_PITCHING_HR = 0.42
  * TEAM_PITCHING_BB = 0.47

To win more games it makes sense the team will need to make fewer errors.  
*Variables with Strongly Negative Correlation with TARGET_WINS:*  
  
There were several batting variables which were related.  
*Positive Correlations between variables*:     
  * TEAM_PITCHING_H and TEAM_BATTING_H = 0.99  
  * TEAM_PITCHING_HR and TEAM_BATTING_HR = 0.99  
  * TEAM_PITCHING_BB and TEAM_BATTING_BB = 0.99    
  * TEAM_PITCHING_SO and TEAM_BATTING_SO = 0.99   
  


## Missing values

```{r, missing-values}


round(100*colSums(is.na(baseball_df))/nrow(baseball_df),2)


```

In terms of missing values, there are two variables missing many obserations. TEAM_BATTING_HBP is missing over 90% of its values,
while TEAM_BASERUN_CS is missing just around 30%. 

```{r, missing-values}

#New DF with Missings Removed
baseball_df_mv <- baseball_df[, !names(baseball_df) %in% c('TEAM_BATTING_HBP','TEAM_BASERUN_CS','TEAM_FIELDING_DP')]
summary(baseball_df_mv)

#Impute NAs with Median
baseball_df_imputed <- mice(baseball_df_mv, m=5, maxit = 5, method = 'pmm')
baseball_df_final <- complete(baseball_df_imputed)
summary(baseball_df_final)

ggplot(melt(baseball_df_final), aes(x=value)) + geom_histogram() + facet_wrap(~variable, scale='free') + labs(x='', y='Frequency')

#Replace Error Maxs
baseball_df_final$TEAM_PITCHING_H[baseball_df_final$TEAM_PITCHING_H > 3*sd(baseball_df_final$TEAM_PITCHING_H)] <- median(baseball_df_final$TEAM_PITCHING_H)
baseball_df_final$TEAM_PITCHING_BB[baseball_df_final$TEAM_PITCHING_BB > 3*sd(baseball_df_final$TEAM_PITCHING_BB)] <- median(baseball_df_final$TEAM_PITCHING_BB)
baseball_df_final$TEAM_PITCHING_SO[baseball_df_final$TEAM_PITCHING_SO > 3*sd(baseball_df_final$TEAM_PITCHING_SO)] <- median(baseball_df_final$TEAM_PITCHING_SO)
baseball_df_final$TEAM_FIELDING_E[baseball_df_final$TEAM_FIELDING_E > 3*sd(baseball_df_final$TEAM_FIELDING_E)] <- median(baseball_df_final$TEAM_FIELDING_E)

summary(baseball_df_final)

```



## Michael Ippolito's code on residuals. Feel free to keep it remvove it as per the next steps

Define function to look at residuals.

```{r}
ResidualAnalysis <- function(lmod) {
  # Plot residuals
  plot(lmod)
  
  # Shapiro test
  #   The null hypothesis is that the residuals are normal.
  #   If the p-value is small, reject the null (i.e., the residuals are not normal).
  st <- shapiro.test(lmod$residuals)
  if (st$p.value <= 0.05) {
    print(paste0("The p-value of ", st$p.value, " is <= 0.05, so reject the null; i.e., the residuals are NOT normal."))
  } else {
    print(paste0("The p-value of ", st$p.value, " is > 0.05, so do not reject the null; i.e., the residuals are normal."))
  }
}
```

## Residual Analysis

Look at residuals.

```{r}
# Residual analysis
ResidualAnalysis(lm1)
```

### **Model Building**

### **Model Selection**

### **References**
* https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html
* https://r-coder.com/correlation-plot-r/



