---
output:
  pdf_document:
    dev: cairo_pdf
    toc: yes
  html_document:
    theme: flatly
    highlight: textmate
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **House Prices**


### CUNY SPS DATA 621

#### GROUP 2: William Aiken, Donald Butler, Michael Ippolito, Bharani Nittala, and Leticia Salazar

$~$

## Abstract:
Use 250 words or less to summarize your problem, methodology, and major outcomes.

[insert text here]

$~$

## Key words:
Select a few key words (up to five) related to your work.

[insert text here]

$~$

## Introduction:
Describe the background and motivation of your problem.

[insert text here]

$~$


## Literature Review:
Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-the- art. Please cite the relevant papers where appropriate.


[insert text here]

$~$

## Methology:
Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

[insert text here]

$~$

## Discussion of Findings:
Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.). Conclude your findings, limitations, and suggest areas for future work.

[insert text here]

$~$

#### Recommendations:

[insert text here]

$~$

#### Limitations:

[insert text here]

$~$

## Conclusion:

[insert text here]

$~$

## References / Bibliography:
Be sure to cite all references used in the report (APA format).

* H, M. Y. (2022, January 12). Housing prices dataset. Kaggle. Retrieved November 28, 2022, from https://www.kaggle.com/datasets/yasserh/housing-prices-dataset 

* [insert text here]

* [insert text here]

$~$

## Appendices:
* Supplemental tables and/or figures.
* R statistical programming code.


[insert graphs, tables, etc here]

```{r code, eval=FALSE}

```


------------------------------------------------------

START OF OUR CODING HERE! WILL NOT SHOW FOR PDF VERSION

### Load Libraries:

These are the libraries used to explore, prepare, analyze and build our models
```{r libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(corrplot)
```

### Load Data:

We have included the original data sets in our GitHub account and read from this location. Since our data set doesn't come with a training and evaluation data sets we will be splitting our data using the 70% - 30% split. Below we are showing the training data set:
```{r, echo=FALSE}
house_prices <- read.csv("https://raw.githubusercontent.com/letisalba/Data_621/master/Final_Project/csv/Housing.csv")

# make this example reproducible
set.seed(1)

# use 70% of data set as training set and 30% as evaluation set
sample <- sample(c(TRUE, FALSE), nrow(house_prices), replace=TRUE, prob=c(0.7,0.3))
dftrain  <- house_prices[sample, ]
dfeval   <- house_prices[!sample, ]

head(dftrain)
```

$~$

### Data Exploration:

Based on this our training data includes 386 records and 13 variables whereas the evaluation data includes 159 records and 13 variables.

Training:
```{r, echo=FALSE}
str(dftrain)
```

$~$

Evaluation:
```{r, echo=FALSE}
str(dfeval)
```

$~$

Using the `summary()` function lets start exploring the training and evaluation data.

Training:
```{r, echo=FALSE}
summary(dftrain)
```

$~$

Evaluation:
```{r, echo=FALSE}
summary(dfeval)
```

#### Price:  
It is important to recognize that this dataset contains homes with prices above 1 million. It is not clear that this is a US dataset, which would indicate that this is for luxury homes and/or high value markets.

```{r price, echo=FALSE, warning=FALSE, error=FALSE}
options(scipen=5)
hist(dftrain$price / 10^6, main = 'Distribution of Price', xlab = 'Price in millions')
```

#### Area:  
The area variable appears to be square footage of the home. We would traditionally expect that increases in area would lead to increases in price.

```{r area, echo=FALSE, warning=FALSE, error=FALSE}
hist(dftrain$area, main = 'Distribution of Area', xlab = 'Area in Square Feet')

```

#### Bedrooms:  
While we expect increases in the number of bedrooms to increase the price, we also realize that at some point there are diminishing returns that an additional bedroom doesn't have as much of an impact. For example, increasing from one to two bedrooms should have significant increase in price, while increasing from four to five, perhaps not so much.

```{r bedrooms, echo=FALSE}
dftrain %>% count(bedrooms)
```

Based on the distribution of the number of Bedrooms, it may be best to categorize these with dummy variables; 2, 3, and 4+.

#### Bathrooms:  
Similar to the number of bedrooms, we would expect that an increase in bathroom count would lead to increases in price. Although similarly, having more than four bathrooms is likely going to lead to smaller increases.

```{r bathrooms, echo=FALSE}
dftrain %>% count(bathrooms)
```

Based on the distribution of the number of bathrooms, it may be best to categorize these with dummy variables; 2, and 3+.

#### Stories:  
Similar to the number of bedrooms and bathrooms, it would seem to make sense to classify homes with 3 or more floors together by introducing dummy variables; 2, and 3+.

```{r stories, echo=FALSE}
dftrain %>% count(stories)
```

#### Parking:  
We are assuming that the parking variable represents the size of a garage. Similar to other variable the increase in price from no garage to a one car garage would be significant, while additional cars would add some lesser value. It would initially seem to make sense to introduce dummy variables; 1, and 2+.

```{r parking, echo=FALSE}
dftrain %>% count(parking)
```

#### Furnishing Status:  
The furnishing status variable is taking on three values; unfurnished, semi-furnished, and furnished. Since we would consider unfurnished as the default state, we will use dummy variables; semi-furnished and furnished.

```{r furnishingstatus, echo=FALSE}
dftrain %>% count(furnishingstatus)
```

#### Main Road:  
The main road variable is yes/no based on the street of the home. We will replace this with a dummy variable.

```{r mainroad, echo=FALSE}
dftrain %>% count(mainroad)
```

#### Guest Room:  
The guest room variable is yes/no based on the home having a guest room. It is unclear from the dataset source if this is in addition to the number of bedrooms, but we would expect houses with a guest room to have a higher price. We will replace this with a dummy variable.

```{r guestroom, echo=FALSE}
dftrain %>% count(guestroom)
```

#### Basement:  
The basement variable is yes/no based on the home having a basement. It is unclear if having a basement or not would lead to an increase in home price, but we will replace this with a dummy variable for analysis.

```{r basement, echo=FALSE}
dftrain %>% count(basement)
```

#### Hot Water Heating:  
Based on the distribution, we assume that the hot water heating variable represents if the house has in-floor heating, rather than forced air. Based on this assumption, we assume that having this feature would lead to higher house price. The variable will be replaced with a dummy variable for analysis.

```{r hotwaterheating, echo=FALSE}
dftrain %>% count(hotwaterheating)
```

#### Air Conditioning:  
The air conditioning variable indicates if the house has central air conditioning. We would expect homes with air conditioning would have a higher price than those without. The variable will be replaced with a dummy variable.

```{r airconditioning, echo=FALSE}
dftrain %>% count(airconditioning)
```

#### Preferential Area:  
The dataset source doesn't specify exactly what this variable represents. We are assuming that this is a yes/no value if the house is in a preferred neighborhood. We would expect houses with a yes to be higher price than those not.

```{r prefarea, echo=FALSE}
dftrain %>% count(prefarea)
```

### Data Preparation:

Based on our exploration, we do not have any blank values in our dataset. 

#### Clean Function:  
We will introduce a clean function to replace our categorical variables with the dummy values. This will also ensure that our test and train datasets are processed in the same way.

```{r clean, echo=FALSE, warning=FALSE, message=FALSE}
fun_clean_df <- function(df) {
  df <- df %>%
    mutate(bed2 = ifelse(bedrooms == 2,1,0)) %>%
    mutate(bed3 = ifelse(bedrooms == 3,1,0)) %>%
    mutate(bed4plus = ifelse(bedrooms >= 4,1,0)) %>% dplyr::select(-bedrooms) %>%
    mutate(bath2 = ifelse(bathrooms == 2,1,0)) %>%
    mutate(bath3plus = ifelse(bathrooms >= 3,1,0)) %>% dplyr::select(-bathrooms) %>%
    mutate(floor2 = ifelse(stories == 2,1,0)) %>%
    mutate(floor3plus = ifelse(stories >= 3,1,0)) %>% dplyr::select(-stories) %>%
    mutate(car1 = ifelse(parking == 1,1,0)) %>%
    mutate(car2plus = ifelse(parking >= 2,1,0)) %>% dplyr::select(-parking) %>%
    mutate(semifurnished = ifelse(furnishingstatus == 'semi-furnished',1,0)) %>%
    mutate(furnished = ifelse(furnishingstatus == 'furnished',1,0)) %>% dplyr::select(-furnishingstatus) %>%
    mutate(mainroad = ifelse(mainroad == 'yes',1,0)) %>%
    mutate(guestroom = ifelse(guestroom == 'yes',1,0)) %>%
    mutate(basement = ifelse(basement == 'yes',1,0)) %>%
    mutate(hotwaterheating = ifelse(hotwaterheating == 'yes',1,0)) %>%
    mutate(ac = ifelse(airconditioning == 'yes',1,0)) %>% dplyr::select(-airconditioning) %>%
    mutate(neighborhood = ifelse(prefarea == 'yes',1,0)) %>% dplyr::select(-prefarea)
}

dftrain_clean <- fun_clean_df(dftrain)

```

#### Coorelation Plot:  

After cleaning the dataset looking at a correlation plot will give us confirmation about our initial examination for the variables.

```{r corplot, warning=FALSE, message=FALSE}
cor_res <- cor(dftrain_clean)

corrplot(cor_res,
         type = "lower",
         order = "original",
         tl.col = "black",
         tl.srt = 50,
         tl.cex = 1)

```

The correlation plot generally confirms our initial expectations for the data. 

$~$

### Model Building:

```{r, echo=FALSE}

```

$~$

### Model Selection:

```{r, echo=FALSE}

```

